Assignment 1
========================================================

## Loading and preprocessing the data

### Load libraries, data, set system

```{r load, message=FALSE, results='hide'}
require(data.table)
require(ggplot2)
require(gridExtra)
require(scales)

activity = data.table(read.csv("activity.csv"))

Sys.setlocale(locale="en_US.UTF-8")
```

### The data.table 'activity'

```{r show data.table}
activity
```

## What is mean total number of steps taken per day?

### Histogram of the total number of steps taken each day

```{r histogram}
steps.summary = activity[, list(steps.per.day=sum(steps, na.rm=T)), by=date]
ggplot(aes(x = steps.per.day), data = steps.summary) +
    geom_histogram(binwidth=2500)
```

### Calculate the mean and median total number of steps taken per day

```{r mean and median of steps taken per day}
activity[, sum(steps, na.rm=T), by=date][,list(mean=mean(V1), median=median(V1))]
```

## What is the average daily activity pattern?

```{r daily activity pattern, fig.width=14}
average.day = activity[, list(steps_mean=mean(steps, na.rm=T)), by=interval]
average.day$time = seq(c(ISOdate(2000,3,20,23,0,0)), by = "5 min", length.out = 288)

ggplot(data=average.day, aes(x=time, y=steps_mean)) +
    geom_line() +
    scale_x_datetime(breaks="2 hours",
                     labels = date_format("%H:%M"))
```

### Average maximum steps time interval

```{r maximum steps interval}
average.day$time = format(seq(c(ISOdate(2000,3,20,23,0,0)), by = "5 min", length.out = 288), "%H:%M")
average.day[order(-steps_mean)][1]
```

## Imputing missing values

### Sum NA's

```{r sum NAs}
sum(is.na(activity))
```

### Impute the mean for that 5-minute interval

```{r impute mean of interval}
activityImputed = activity
activityImputed = merge(activityImputed, average.day, by=c('interval'))
activityImputed[is.na(steps)]$steps = as.integer(round(activityImputed[is.na(steps)]$steps_mean))
activityImputed
```

### Histogram of the total number of steps taken each day - imputed dataset

```{r histogram imputed data}
steps.summary = activityImputed[, list(steps.per.day=sum(steps, na.rm=T)), by=date]
ggplot(aes(x = steps.per.day), data = steps.summary) +
    geom_histogram(binwidth=2500)
```

### Calculate the mean and median total number of steps taken per day - imputed dataset

```{r mean and median of steps taken per day imputed data}
activityImputed[, sum(steps, na.rm=T), by=date][,list(mean=mean(V1), median=median(V1))]
```

## Differences in activity patterns between weekdays and weekends

```{r Differences in activity patterns between weekdays and weekends, fig.width=14}
# Add new variables: weekDay and weekPart
activityImputed$weekDay = weekdays(as.Date(activityImputed$date))
activityImputed$weekPart = "weekday"
setkey(activityImputed, weekDay)
activityImputed[J(c('Saturday','Sunday')), weekPart:="weekend"]

# Create final dataset
# summarise weekday and weekend
activityPatterns = activityImputed[,list(steps_mean= mean(steps)), by=c("interval","weekPart")]
# convert 'interval' to time serie
activityPatterns$time = seq(c(ISOdate(2000,3,20,23,0,0)), by = "5 min", length.out = 288)

# Create plot of final dataset
ggplot(data=activityPatterns, aes(x=time, y=steps_mean)) +
    geom_line() + ggtitle("Activity Patterns") +
    scale_x_datetime(breaks="2 hours",
                     labels = date_format("%H:%M")) +
    facet_grid(weekPart ~ .)

```